Systematic studies of correlations between different order flow harmonics in $\PbPb$ collisions at $\snn = 2.76$~TeV

### 2017-05-30 longSC.tex
sudhir comment1

\section{Introduction}
\label{sec:intro}
The main emphasis of the ultra-relativistic heavy-ion collision programs at the Relativistic Heavy Ion Collider (RHIC) and the Large Hadron Collider (LHC) is to study the deconfined phase of strongly interacting nuclear matter, the Quark-Gluon Plasma (QGP). 
%This matter exhibits strong anisotropic flow in the plane transverse to the beam direction, which is driven by anisotropic pressure gradients, resulting in more particles emitted in the direction of the largest gradients.
The matter produced in a heavy-ion collision exhibits strong collective radial expansion. Due to anisotropic pressure gradients in the plane transverse to the beam direction, more particles emitted in the direction of the largest gradients result in anisotropic transverse flow.
The large elliptic flow discovered at RHIC energies~\cite{Ackermann:2000tr} is also observed at LHC energies~\cite{Aamodt:2010pa,Adam:2016izf}. The measurements are well described by calculations utilizing viscous hydrodynamics~\cite{Romatschke:2007mq,Shen:2011eg,Schenke:2011zz,Bozek:2012qs,Gale:2012rq,Hirano:2010je}.
%and microscopic transport models ~\cite{Xu:2007jv}.  
These calculations also demonstrated that the shear viscosity to the entropy density ratio ($\eta/s$) of the QGP in heavy-ion collisions at RHIC and LHC energies is close to a universal lower bound $1/4\pi$~\cite{Kovtun:2004de}.

@@

An important input to the hydrodynamic model simulations is the initial distribution of energy density in the transverse plane (the initial density profile), which is usually estimated from the probability distribution of nucleons in the incoming nuclei.
This initial energy density profile can be quantified by calculating the distribution of the spatial eccentricities $\epsilon_n$~\cite{Alver:2010gr},
\begin{equation}
 \varepsilon_{n} e^{in\Phi_{n}} = -\{r^n e^{in\phi}\}/ \{r^n\},
  \label{eq:eccentricities}
\end{equation}
where the curly brackets denote the average over the transverse plane, i.e. $\{\cdots\}$ = $\int dxdy$\, $e(x,y,\tau_0)$ $(\cdots)$, $r$ is the distance to the system's center of mass, $e(x,y,\tau_0)$ is the energy density at the initial time $\tau_0$, and $\Phi_{n}$ is the participant plane angle (see Refs.~\cite{Teaney:2010vd,Niemi:2015qia}).
There is experimental and theoretical evidence~\cite{Alver:2010gr,Qiu:2011iv,ALICE:2011ab} that the lower order harmonics, $v_2$ and $v_3$, are to a good approximation linearly proportional to the deformations in the initial energy density in the transverse plane (e.g. $v_n \propto \varepsilon_n$ for $n~=~$2 or 3).
Harmonic $v_4$ and higher order flow harmonics can arise from initial anisotropies in the same harmonic~\cite{Alver:2010gr,Teaney:2010vd,Gubser:2010ui,Hatta:2014jva} (linear response) or can be induced by lower order harmonics~\cite{Bravina:2013xla,Bravina:2013ora} (nonlinear response).
Therefore, the higher harmonics ($n>3$) can be understood as superpositions of linear and nonlinear responses, through which they are correlated with lower order harmonics~\cite{Teaney:2012ke,Bravina:2013ora,Gubser:2010ui,Hatta:2014jva}. When the order of the harmonic is large, the nonlinear response contribution in viscous hydrodynamics is dominant and increases in more peripheral collisions~\cite{Teaney:2012ke,Bravina:2013ora}.
The magnitudes of the viscous corrections as a function of $p_{\rm T}$ for $v_4$ and $v_5$ are sensitive to the ansatz used for the viscous distribution function, and to the correction for the equilibrium distribution at hadronic freeze-out~\cite{Luzum:2010ad,Teaney:2012ke}.
%and the $p_{\rm T}$ dependence of linear and nonlinear response contributions to the higher order flow harmonics is discussed further in~\cite{Teaney:2012ke}. 
Hence, studies of the correlations between higher order ($n>3$) and lower order ($v_2$ or $v_3$) harmonics and their $p_{\rm T}$ dependence can help to understand the viscous correction to the momentum distribution at hadronic freeze-out which is among the least understood parts of hydrodynamic calculations~\cite{Dusling:2009df,Teaney:2012ke,Molnar:2014fva,Niemi:2015qia}.

@@

For reasons discussed in \cite{ALICE:2016kpq,Bilandzic:2013kga}, the correlations between different flow harmonics cannot be studied experimentally with the same set of observables introduced in \cite{Niemi:2012aj}. 
Based on \cite{Bilandzic:2013kga}, new flow observables obtained from multiparticle correlations, \textit{Symmetric Cumulants}~(SC), were introduced. 
%The SC observables are nearly insensitive to nonflow and quantify the correlation of the amplitudes of two different flow harmonics.
%The first measurements of SC observables were recently published by the ALICE Collaboration in \cite{ALICE:2016kpq}.

@@

Another large contribution to the systematic uncertainty originates from azimuthal non-uniformities in the reconstruction efficiency. In order to estimate its effects, we use the AMPT model (see Sec.~\ref{sec:theory}) which has a uniform distribution in azimuthal angle.
Detector inefficiencies were introduced to mimic the non-uniform azimuthal distribution in the data. For the observables SC(5,2), SC(5,3) and SC(4,3) the variation due to non-uniform acceptance is about 9\%, 17\% and 11\%, respectively.
Overall, the systematic uncertainties are larger for SC(5,3) and SC(5,2) than for the lower harmonics of SC$(m,n)$.
%This is because $v_{n}$ decreases with increasing $n$ and becomes more sensitive to azimuthal modulation due to detector imperfections. 
%Furthermore, the various AMPT models (see the details in Sec.~\ref{sec:theory}) were used to check the sensitivities of the observables due to the strength of the signals we are measuring, it turned out that the effect is negligible.

@@

Unlike SC$(m,n)$, the NSC$(m,n)$ results with the higher order flow harmonics show almost the same order of the correlation strength as the lower order flow harmonic correlations (NSC(3,2) or NSC(4,2)).
This demonstrates the advantage of using the normalized SC observables in which the correlation strength between flow harmonics is not hindered by the differences in magnitudes of different flow harmonics. The NSC(4,3) magnitude is comparable to NSC(3,2) and one finds that a hierarchy, NSC(5,3) $>$ NSC(4,2) $>$ NSC(5,2), holds for centrality ranges $>$~20\% within the errors as shown in Fig.~\ref{fig:Figure_1}(b).
These results indicate that the lower order harmonic correlations are larger than higher order harmonic correlations, not only because of the correlation strength itself but also because of the strength of the individual flow harmonics. 
The SC(5,2) magnitude is larger than SC(5,3), but the normalized correlation between $v_5$ and $v_3$ is stronger than the normalized correlation between $v_5$ and $v_2$.

>>>>

\section{Introduction}
\label{sec:intro}
The main emphasis of the ultra-relativistic heavy-ion collision program at the Relativistic Heavy Ion Collider (RHIC) and the Large Hadron Collider (LHC) is to study the deconfined phase of strongly interacting nuclear matter, the Quark-Gluon Plasma (QGP). 
%This matter exhibits strong anisotropic flow in the plane transverse to the beam direction, which is driven by anisotropic pressure gradients, resulting in more particles emitted in the direction of the largest gradients.
The matter produced in a heavy-ion collision exhibits strong collective radial expansion. Due to anisotropic pressure gradients in the plane transverse to the beam direction, more particles emitted in the direction of the largest gradients result in anisotropic transverse flow.
The large elliptic flow discovered at RHIC energies~\cite{Ackermann:2000tr} is also observed at LHC energies~\cite{Aamodt:2010pa,Adam:2016izf}. The measurements are well described by calculations utilizing viscous hydrodynamics~\cite{Romatschke:2007mq,Shen:2011eg,Schenke:2011zz,Bozek:2012qs,Gale:2012rq,Hirano:2010je}.
%and microscopic transport models ~\cite{Xu:2007jv}.  
These calculations also demonstrated that the shear viscosity to the entropy density ratio ($\eta/s$) of the QGP in heavy-ion collisions at RHIC and LHC energies is close to a universal lower bound $1/4\pi$~\cite{Kovtun:2004de}.

@@

An important input to the hydrodynamic model simulations is the initial distribution of energy density in the transverse plane (the initial density profile), which is usually estimated from the probability distribution of nucleons in the incoming nuclei.
This initial energy density profile can be quantified by calculating the distribution of the spatial eccentricities $\epsilon_n$~\cite{Alver:2010gr},
\begin{equation}
 \varepsilon_{n} e^{in\Phi_{n}} = -\{r^n e^{in\phi}\}/ \{r^n\},
  \label{eq:eccentricities}
\end{equation}
where the curly brackets denote the average over the transverse plane, i.e. $\{\cdots\}$ = $\int dxdy$\, $e(x,y,\tau_0)$ $(\cdots)$, $r$ is the distance to the system's center of mass, $e(x,y,\tau_0)$ is the energy density at the initial time $\tau_0$, and $\Phi_{n}$ is the participant plane angle (see Refs.~\cite{Teaney:2010vd,Niemi:2015qia}).
There is experimental and theoretical evidence~\cite{Alver:2010gr,Qiu:2011iv,ALICE:2011ab} that the lower order harmonics, $v_2$ and $v_3$, to a good approximation, are linearly proportional to the deformations in the initial energy density in the transverse plane (e.g. $v_n \propto \varepsilon_n$ for $n~=~$2 or 3).
Harmonic $v_4$ and higher order flow harmonics can arise from initial anisotropies in the same harmonic~\cite{Alver:2010gr,Teaney:2010vd,Gubser:2010ui,Hatta:2014jva} (linear response) or can be induced by lower order harmonics~\cite{Bravina:2013xla,Bravina:2013ora} (nonlinear response).
Therefore, the higher harmonics ($n>3$) can be understood as superpositions of linear and nonlinear responses, through which they are correlated with lower order harmonics~\cite{Teaney:2012ke,Bravina:2013ora,Gubser:2010ui,Hatta:2014jva}. When the order of the harmonic is large, the nonlinear response contribution in viscous hydrodynamics is dominant and increases in more peripheral collisions~\cite{Teaney:2012ke,Bravina:2013ora}.
The magnitudes of the viscous corrections as a function of $p_{\rm T}$ for $v_4$ and $v_5$ are sensitive to the ansatz used for the viscous distribution function, and to the correction for the equilibrium distribution at hadronic freeze-out~\cite{Luzum:2010ad,Teaney:2012ke}.
%and the $p_{\rm T}$ dependence of linear and nonlinear response contributions to the higher order flow harmonics is discussed further in~\cite{Teaney:2012ke}. 
Hence, studies of the correlations between higher order ($n>3$) and lower order ($v_2$ or $v_3$) harmonics and their $p_{\rm T}$ dependence can help to understand the viscous correction to the momentum distribution at hadronic freeze-out which is among the least understood parts of hydrodynamic calculations~\cite{Dusling:2009df,Teaney:2012ke,Molnar:2014fva,Niemi:2015qia}.

@@

For reasons discussed in \cite{ALICE:2016kpq,Bilandzic:2013kga}, the correlations between different flow harmonics cannot be studied experimentally with the set of observables introduced in \cite{Niemi:2012aj}. 
Based on \cite{Bilandzic:2013kga}, new flow observables obtained from multiparticle correlations, \textit{Symmetric Cumulants}~(SC), were introduced. 
%The SC observables are nearly insensitive to nonflow and quantify the correlation of the amplitudes of two different flow harmonics.
%The first measurements of SC observables were recently published by the ALICE Collaboration in \cite{ALICE:2016kpq}.

@@

Another large contribution to the systematic uncertainty originates from azimuthal non-uniformities in the reconstruction efficiency. In order to estimate its effects, we use the AMPT model (see Sec.~\ref{sec:theory}) which has a uniform distribution in azimuthal angle.
Detector inefficiencies were introduced to mimic the non-uniform azimuthal distribution in the data. For the observables SC(5,2), SC(5,3) and SC(4,3) the variation due to non-uniform acceptance is about 9\%, 17\% and 11\%, respectively.
Overall, the systematic uncertainties are larger for SC(5,3) and SC(5,2) than for the lower harmonics of SC$(m,n)$.
This is because $v_{n}$ decreases with increasing $n$ and becomes more sensitive to azimuthal modulation due to detector imperfections. 
%Furthermore, the various AMPT models (see the details in Sec.~\ref{sec:theory}) were used to check the sensitivities of the observables due to the strength of the signals we are measuring, it turned out that the effect is negligible.

@@

Unlike SC$(m,n)$, the NSC$(m,n)$ results with the higher order flow harmonics show almost the same order of the correlation strength as the lower order flow harmonic correlations (NSC(3,2) or NSC(4,2)).
This demonstrates the advantage of using the normalized SC observables in which the correlation strength between flow harmonics is not hindered by the differences in magnitudes of different flow harmonics. The NSC(4,3) magnitude is comparable to NSC(3,2) and one finds that a hierarchy, NSC(5,3) $>$ NSC(4,2) $>$ NSC(5,2), holds for centrality ranges $>$~20\% within the errors as shown in Fig.~\ref{fig:Figure_1}(b).
The SC(5,2) magnitude is larger than SC(5,3), but the normalized correlation between $v_5$ and $v_3$ is stronger than the normalized correlation between $v_5$ and $v_2$. 
These results indicate that the lower order harmonic correlations are larger than higher order harmonic correlations, not only because of the correlation strength itself but also because of the strength of the individual flow harmonics.
